\documentclass[10pt]{memoir}

% based on kieran healy's memoir modifications
\usepackage{mako-mem}
\chapterstyle{article-2}
\pagestyle{mako-mem}

\usepackage{ucs}
\usepackage[utf8x]{inputenc}

%\usepackage{kpfonts}
%\usepackage[bitstream-charter]{mathdesign}
\usepackage{fbb}
\usepackage[T1]{fontenc}
%\usepackage{textcomp}

%\renewcommand{\rmdefault}{ugm}
%\renewcommand{\sfdefault}{phv}
%

\usepackage[letterpaper,left=1.25in,right=1.25in,top=1.25in,bottom=1.25in]{geometry}

% packages i use in essentially every document
\usepackage{graphicx}
\usepackage{enumerate}

% Setup list environments
\usepackage{enumitem}
\setlist[description]{
  topsep=0pt,
  before=\vspace{5pt},
  after=\vspace{0pt},
  itemsep=2pt,
  labelsep=0pt
}

% packages i use in many documents but leave off by default
% \usepackage{amsmath, amsthm, amssymb}
% \usepackage{dcolumn}
% \usepackage{endfloat}

% Set paragraph indents and spacing
\setlength{\parindent}{0pt}
\setlength{\parskip}{.5\baselineskip}
%\usepackage[document]{ragged2e}

% adjust section title formatting
\usepackage{titlesec}
\titlespacing\section{-8pt}{8pt plus 2pt minus 2pt}{-6pt}
\titlespacing\subsection{-8pt}{8pt plus 2pt minus 2pt}{-6pt}
\titlespacing\subsubsection{0pt}{8pt plus 2pt minus 2pt}{10pt}

% allows full, in-line citations
\usepackage{bibentry} 

% add bibliographic stuff 
\usepackage[round, numbers]{natbib} \def\citepos#1{\citeauthor{#1}'s (\citeyear{#1})} \def\citespos#1{\citeauthor{#1}' (\citeyear{#1})}
\renewcommand{\bibnumfmt}[1]{}

% define colors from http://www.colorado.edu/brand/visual-identity/typography-color
\usepackage[usenames,dvipsnames]{color}
\definecolor{CUGold}{RGB}{207,184,124}
\definecolor{CUDarkGray}{RGB}{86,90,92}
\definecolor{CULightGray}{RGB}{162,164,163}

% customize URLs
%\usepackage{url}
%\usepackage{breakurl} 
\usepackage[breaklinks, bookmarks, bookmarksopen]{hyperref}
\hypersetup{colorlinks=true, linkcolor=Blue, citecolor=Black, filecolor=Blue,
    urlcolor=Blue, unicode=true, breaklinks=true}

% create a "reading list" environment to format the items
\newenvironment{readinglist}{
\begin{list}{}{\leftmargin=8pt \itemindent=0em}
  \setlength{\itemsep}{8pt}
  \setlength{\parskip}{0em}
  \setlength{\parsep}{1em}
  \setlength{\parindent}{8em}}
{\end{list}}

% Course/Instructor metadata -- alter as neded
\def\myauthor{Brian Keegan}
\def\mycoursename{Peer Production and Crowdsourcing}
\def\mycourselisting{INFO 3501/5501}
\def\myoffice{ENVD 201}
\def\myclassroom{Hellems 137}
\def\mymeetingtime{Monday \& Wednesday 17:30--18:45}
\def\mydate{Fall 2016}
\def\myemail{brian.keegan@colorado.edu}
\def\myweb{http://www.brianckeegan.org}
\def\myofficehours{Monday 15:00--17:00 or by appointment}

% Some that I'm not using here:
\def\mytitle{Assistant Professor }
\def\mycopyright{\myauthor}
\def\myphone{(+1) 617-803-6971}
\def\mystreet{1060 18th St.}
\def\mycity{Boulder, CO 80302}

\begin{document}

\nobibliography*

%\baselineskip 14.2pt

\title{
    \textit{\normalsize{\textcolor{CUGold}{\textbf{Problems in Information Science:}}}}\\
    \textbf{\huge{\mycoursename}}\\
    \vspace{5pt} \normalsize{\mycourselisting}
    }

\author{\mydate\\ \mymeetingtime\\ \myclassroom}

\date{\normalsize{\mytitle \myauthor\\
       E-mail: \href{mailto:\myemail}{\myemail}\\
       Office: \myoffice\\
       Office hours: \myofficehours}}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%
%% Acknowledgements %%
%%%%%%%%%%%%%%%%%%%%%%
% This syllabus template was made in LaTeX by Brian Keegan and is distributed as Free Software under the GNU GPL v3. It was built using style templates created by Aaron Shaw, Benjamin Mako Hill, and Kieran Healy.

\section{\textbf{Course Description}}

This is a research seminar that will analyze the social and technical mechanisms that enable popular peer production and crowdsourcing systems like Wikipedia. Understanding how social processes and technical structures intersect to enable new kinds of interaction is a central question within information science. How does the design of a peer production system tap into basic social, psychological, and organizational processes? What kinds of behavioral data can researcher extract and analyze from these systems? How can the success or failure of these systems inform the design of alternative economic and political models? This course is part of the ``Problems in Information Science'' series, which brings contemporary research to the classroom in the form of project-based investigation.

\subsection{Learning Objectives and Course Design}

By the end of the semester, students should be able to: (1) discuss and compare social processes occurring across different peer production and crowdsourcing systems; (2) identify and collect data generated by these systems; (3) design exploratory research using these data and interpret their findings; and (4) propose and evaluate the feasibility of peer production and crowdsourcing designs to expand to other social, cultural, economic, and political spheres.

The course consists of three, five-week units. The first unit focuses on peer production and crowdsourcing as \textit{social} systems by exploring how motivations, norms, and governance interact. The second unit focuses on peer production and crowdsourcing as \textit{technical} systems by exploring the systems that structure user interactions and the kinds of data they record. The third unit focuses on peer production and crowdsourcing as \textit{alternative} systems to prevailing models for decision making, collaboration, and economic production.

Class will meet twice per week on Monday and Wednesday from 17:30 to 18:45 (5:30pm to 6:45pm) in \myclassroom. The format of each class will vary between lecture, discussion, activities, and lab formats depending on the learning objectives of that week. Students will complete three kinds of assignments over the course of the class: (1) weekly reflections, (2) lab exercises, and (3) a final project.

\subsection{Prerequisites}
There are no formal prerequisites for this course. However, you should have some familiarity (prior class, research project, internship experience, MOOC coursework, \textit{etc.}) with methods for data mining (web scraping, information retrieval from APIs, \textit{etc}.), quantitative analysis (statistical inference, machine learning, \textit{etc}.), and data visualization. Knowledge of statistical software or scientific computing languages such as \texttt{R} or \texttt{Python} will also be helpful, but is not required. If you have questions or concerns about these prerequisites, please \href{mailto:brian.keegan@colorado.edu}{email me}.

\subsection{Requirements}
As a research seminar, students' regular and sustained participation in all class activities is essential. Attending and participating in all seminar meetings and completing all assignments. If you need to be excused from attendance or need an extension, please \href{mailto:brian.keegan@colorado.edu}{notify me via email} at least 24 hours in advance.

\subsection{Course Website and Materials}
There is no textbook for the class and we will read magazine articles, academic papers, or book chapters. For many of the class sessions I have also included a sub-section of ``Additional Readings'' that are purely optional but may be useful if students are interested in reading in further depth to support a research project, preparing for prelims, \textit{etc}. Materials for the class (including a ``live'' version of the course schedule, readings, discussion board, lab assignments, tutorials, and data) will be made available through \mbox{Desire2Learn}:
\vspace{-8pt}
    \begin{center}
    \href{https://https://learn.colorado.edu/d2l/home/179594}{https://learn.colorado.edu/d2l/home/179594}
    \end{center}
\vspace{-8pt}
Once the semester begins, this PDF version of the syllabus will be revised infrequently and any revised requirements will be posted as announcements and updated course schedule to \mbox{Desire2Learn}.

For students interested in understanding Wikipedia in greater depth, I would recommend the following books:

\begin{readinglist} 
    \item \bibentry{ayers_wikipedia_2008}
    \item \bibentry{benkler_wealth_2006}
    \item \bibentry{jemielniak_common_2014}
    \item \bibentry{kraut_building_2012}
    \item \bibentry{reagle_good_2010}
\end{readinglist}

\subsection{Assignments}
There are three kinds of assignments in the class: (1) reading responses, (2) lab assignments, and a (3) final project. Assignments should be submitted to me via \mbox{Desire2Learn}. I strongly prefer submissions to be in PDF format to maximize compatibility and reduce security risks. Students will complete and submit their work individually but if they work with others, they should acknowledge their collaborators by name on the submission.

\begin{description}[itemsep=0pt,labelsep=0pt]
    \item[Readings Responses] -- For weeks that do not have lab assignments, students will write at least a 500-word response to the readings for that week. Reading responses should be submitted before 9:00 (am) on their due dates for me to review them before class to support discussion. There are 8 reading responses assigned over the course of the semester. The schedule in the syllabus below has the suggested prompts but students should feel free (1) to surface and respond to other themes from the week's readings, (2) to juxtapose the readings to prior weeks' topics, or (3) to review one of the papers listed under Additional Readings. If students are struggling with formulating a response, consider using this \href{http://www.brianckeegan.com/2014/03/checklist-for-reviewing-and-thus-writing-a-research-paper/}{Checklist for Reviewing a Research Paper}.
    \item[Lab Assignments ] -- Lab assignments are intended to develop students' skill and confidence performing exploratory data analyses using real world data from peer production systems. There are 5 lab assignments assigned over the course of the semester. We will work through the lab assignments during one class session but additional work may be required outside of class to write up discussions of the results. The majority of these assignments will be framed around the English Wikipedia, but students should feel free to extend these analyses to other wikis (other language editions, Wikia wikis, \textit{etc}.). Templates for completing the lab assignments will be provided in Jupyter Notebooks to enable students with minimal coding experience to complete them. 
    \item[Final Project ] -- In lieu of having a final exam, students will submit a project that (1) dives deeper into a topical area, (2) extends the lab assignments into a larger data analysis, (3) critically analyzes an existing peer production or crowdsourcing system, or (4) develops a detailed proposal for applying a peer production or crowdsourcing model to an untapped domain. The final project will be due at the end of the semester.
\end{description}

\subsection{Statistical Computing}
You will need to use statistical computing software in this class. You are welcome to use any scripting or statistical software you like (Python, \texttt{R}, \textit{etc}.); however, I will use \href{http://jupyter.org/}{Jupyter notebooks} written in Python for all in-class examples and solutions to the problem sets. I recommend using the \href{https://www.continuum.io/why-anaconda}{Anaconda distribution} of Python that includes the vast majority of libraries and functionality to support data collection and analysis for this class. You are however welcome to use any statistical software with which you are comfortable to complete lab assignments.

\subsection{Evaluation} 
Your final grade for the course will be based on my evaluation of the three kinds of assignments as well as your general participation within discussions during class and online. 

    \begin{description}[itemsep=0pt,labelsep=0pt]
        \item[Reading Responses]~($30\%$) Responses that provide integrate other theories or perspectives, identify methodological limitations, or persuasively discuss implications in the context of contemporary topics will be evaluated more strongly than responses that simply summarize the readings or only make superficial connections among the themes. 
        \item[Lab Assignments]~($30\%$) Students should include their well-commented code documenting your process as well as the final outputs (\textit{e.g.}, submitting the Jupyter Notebook file). Students should be prepared to explain your implementation in a 1-on-1 code review if I request it. Assignments will be primarily evaluated on their completeness, use of good documentation practice, and clarity when discussing results.
        \item[Final Project]~($25\%$) Criteria for proposing and evaluating these final projects will be discussed later in the course. 
        \item[Participation]~($15\%$) Students demonstrating consistent attendance, active engagement, high preparation, deep curiosity, sustained retention, positive attitude, persuasive arguments, and well-formatted assignments will earn the most credit.
    \end{description}

\section{Course Policies}

\subsection{In-Class Confidentiality}
The success of this class depends on participants feeling comfortable sharing questions, ideas, concerns, and confusions about works-in-progress, the research process, and their personal experiences. These assignments and discussions should be considered confidential and generally should not be discussed outside of class. You may read and comment on classmates' writing, code, and images for the sole purpose of use within this class. You may not use, run, copy, perform, display, distribute, modify, translate, or create derivative works of other students' work outside of this class without their expressed written consent or formal license. Furthermore, you may not create any audio, video, or other records during class time nor may you publicly share comments attributable to other people's identities without that person's permission.

\subsection{Faculty Interaction}
In addition to teaching this class, I also (1) manage a research program; (2) advise students; (3) perform service for the academic community; and (4) live my life as a private citizen. I will check e-mail between 8:00 and 18:00 on non-holiday business days and try to respond within 24 hours. I welcome online or offline interactions outside of class, however these are not appropriate spaces for discussing class matters. \href{maito:brian.keegan@colorado.edu}{E-mailing me}, coming to my office hours, or scheduling an appointment are the best ways to ask questions, discuss concerns, or get feedback outside of class.

\subsection{Research Ethics and Professional Conduct}
This course will involve participating in online communities, analyzing data available on the web, and prototyping alternative peer production and crowdsourcing systems. At all times you should make sure that your actions simultaneously minimize the risk of harm to other people as well to yourself. You should follow the Terms of Service and Privacy Policy when collecting data from websites; only collect publicly-accessible data; identify yourself as a student or researcher when interacting with others; and allow participants to remove themselves from your study if they request it. More details about research ethics with online data can be found in the Association for Computing Machinery's \href{https://www.acm.org/about-acm/acm-code-of-ethics-and-professional-conduct}{Code of Ethics and Professional Conduct}, Association for Internet Researchers' \href{http://aoir.org/ethics/}{ethics working committee reports}, American Psychological Association's \href{http://www.apa.org/science/leadership/bsa/internet/internet-report.aspx}{online research report}.

\subsection{Deadlines and Absences}
If something causes you to miss a deadline or a class, please contact me. If you request --- and obtain --- an incomplete for the course and/or an extension on the final project (note: I \textit{strongly} discourage this!), please allow at least 4 weeks after you submit your completed work for me to submit a grade. Keep this in mind if you will need the grade in order to receive your fellowship/diploma/visa/\textit{etc}. by a particular date.

\subsection{Accommodations for Disabilities}
I am committed to providing everyone the support and services needed to participate in this course. If you qualify for accommodations because of a disability, please submit to your professor a letter from Disability Services in a timely manner (for exam accommodations provide your letter at least one week prior to the exam) so that your needs can be addressed. Disability Services determines accommodations based on documented disabilities. Contact Disability Services at 303-492-8671 or by e-mail at \href{mailto:dsinfo@colorado.edu}{dsinfo@colorado.edu}. If you have a temporary medical condition or injury, see \href{http://www.colorado.edu/disabilityservices/students/temporary-medical-conditions}{Temporary Medical Conditions} guidelines under Quick Links at \href{http://www.colorado.edu/disabilityservices/students}{Disability Services} website and discuss your needs with me.

\subsection{Religious Observance}
Campus policy regarding \href{http://www.colorado.edu/policies/observance-religious-holidays-and-absences-classes-andor-exams}{religious observances} requires that faculty make every effort to deal reasonably and fairly with all students who, because of religious obligations, have conflicts with scheduled exams, assignments or required assignments/attendance. If this applies to you, please email me directly as soon as possible at the beginning of the term. 

\subsection{Classroom Behavior}
Students and faculty each have responsibility for maintaining an appropriate learning environment. Those who fail to adhere to such behavioral standards may be subject to discipline. Professional courtesy and sensitivity are especially important with respect to individuals and topics dealing with differences of race, color, culture, religion, creed, politics, veteran’s status, sexual orientation, gender, gender identity and gender expression, age, ability, and nationality.  Class rosters are provided to the instructor with the student's legal name. I will gladly honor your request to address you by an alternate name or gender pronoun. Please advise me of this preference early in the semester so that I may make appropriate changes. For more information, see the policies on \href{http://www.colorado.edu/policies/student-classroom-and-course-related-behavior}{class behavior} and the \href{http://www.colorado.edu/osc/#student_code}{student code}.

\subsection{Harassment and Discrimination}
The University of Colorado Boulder (CU Boulder) is committed to maintaining a positive learning, working, and living environment. CU Boulder will not tolerate acts of sexual misconduct, discrimination, harassment or related retaliation against or by any employee or student. CU's \href{http://www.colorado.edu/policies/discrimination-and-harassment-policy-and-procedures}{Sexual Misconduct Policy} prohibits sexual assault, sexual exploitation, sexual harassment, intimate partner abuse (dating or domestic violence), stalking or related retaliation. CU Boulder's \href{http://www.colorado.edu/policies/discrimination-and-harassment-policy-and-procedures}{Discrimination and Harassment Policy} prohibits discrimination, harassment or related retaliation based on race, color, national origin, sex, pregnancy, age, disability, creed, religion, sexual orientation, gender identity, gender expression, veteran status, political affiliation or political philosophy. Individuals who believe they have been subject to misconduct under either policy should contact the Office of Institutional Equity and Compliance (OIEC) at 303-492-2127. Information about the OIEC, the above referenced policies, and the campus resources available to assist individuals regarding sexual misconduct, discrimination, harassment or related retaliation can be found at the \href{http://www.colorado.edu/institutionalequity/}{OIEC website}.

\subsection{Honor Code}
All students enrolled in a University of Colorado Boulder course are responsible for knowing and adhering to the \href{http://www.colorado.edu/policies/academic-integrity-policy}{academic integrity policy} of the institution. Violations of the policy may include: plagiarism, cheating, fabrication, lying, bribery, threat, unauthorized access, clicker fraud, resubmission, and aiding academic dishonesty. All incidents of academic misconduct will be reported to the Honor Code Council (\href{mailto:honor@colorado.edu}{honor@colorado.edu}; 303-735-2273). Students who are found responsible for violating the academic integrity policy will be subject to nonacademic sanctions from the Honor Code Council as well as academic sanctions from the faculty member. Additional information regarding the academic integrity policy can be found at \href{http://honorcode.colorado.edu}{honorcode.colorado.edu}. 

\section{Acknowledgements}
The design and format of this course borrows from similar courses offered by other professors whose research and thinking has been very influential on me: \href{http://aaronshaw.org/}{Aaron Shaw}'s ``\href{http://aaronshaw.org/teaching/2014/occ/syllabus/Shaw-occ-syllabus2014.pdf}{Online Communities \& Crowds}'' at Northwestern University's School of Communication, \href{https://mako.cc/academic/}{Benjamin Mako Hill}'s ``\href{http://wiki.communitydata.cc/Interpersonal_Media_(Fall_2015)}{Interpersonal Media}'' at the University of Washington's Department of Communication, and \href{http://reagle.org/joseph}{Joseph Reagle}'s ``\href{http://reagle.org/joseph/2014/oc/oc-syllabus-FA.html}{Online Communities}'' at Northeastern University's Department of Communication Studies.

%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%
%%% COURSE OUTLINE %%%
%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%

\newpage
\section{\textbf{Course Outline}}

As described above, all readings and assigned tasks are due \emph{prior} to the class alongside which they are listed. Items listed in the ``Additional Readings'' subsection for each week are not required. I may adjust the list of readings or the schedule as needed throughout the quarter, so please consult the online version of the schedule for the most up-to-date information. We will meet in class a total of 29 times over the course of the the 16-week semester.

%%%%%%%%%%%%%
%% WEEK 01 %%
%%%%%%%%%%%%%

% Confirm research computing environments for Python and \texttt{R} are properly configured and ready for data collection and analysis. 

\section{0 -- Introductions \& Research Ethics}
\textcolor{CUGold}{\textbf{Monday, August 22}}\\
Review syllabus, learning objectives, and course policies. Introduce yourselves and share motivations and goals for the class. Discuss ethical considerations when conducting online research such as collecting ``public'' data, following Terms of Service, and protecting human subjects.

    \subsection{Readings}
    \begin{readinglist}
        \item This syllabus.
        \item \bibentry{bruckman_teaching_2006}
        \item \bibentry{zimmmer_okcupid_2016}
        \item \bibentry{wiki_terms_of_use}
    \end{readinglist}

\section{1 -- A Wager}
\textcolor{CUGold}{\textbf{Wednesday, August 24}}\\
In 2006, Nick Carr and Yochai Benker made a wager about the future of online information production. More than a decade later, who --- if anyone --- won? We will develop preliminary definitions of peer production and crowdsourcing and assess their fit to the current landscape.

    \subsection{Readings}
    \begin{readinglist}
        \item \bibentry{carr_calacaniss_2006a}
        \item \bibentry{carr_benkler_2006b}
        \item \bibentry{carr_pay_2012}
        \item \bibentry{benkler_carr-benkler_2012}
    \end{readinglist}

    \subsection{Assignments}
    \begin{description}%[itemsep=2pt,labelsep=0pt]
        \item[Reading Response 1 ] -- due Wednesday, August 24. ``Will the decentralized web collectives be able to operate successfully outside `the price system' and without `managerial structure'?'' Do the most influential web platforms in 2016 follow peer production processes or are they price-incentivized systems? Given how web platforms now operate, what considerations did Carr and Benkler overlook?
    \end{description}

    \subsection{Additional Readings}
    \begin{readinglist}
        \item \bibentry{benkler_ch3_2006}
        \item \bibentry{carr_amorality_2005}
        \item \bibentry{keen_web_2006}
        \item \bibentry{lanier_digital_2006}
    \end{readinglist}

%%%%%%%%%%%%%
%% WEEK 02 %%
%%%%%%%%%%%%%
\section{2 -- Motivation}
\textcolor{CUGold}{\textbf{Monday, August 29}}\\
What motivates participants to join and contribute to online communities? We will examine how different motivations contribute to or detract from the success of peer production systems. 

    \subsection{Readings}
    \begin{readinglist}
        \item \bibentry{wikipedia_why_2016}
        \item \bibentry{kraut_building_2012}. Chapter 2.
        \item \bibentry{nov_what_2007}
    \end{readinglist}
    
    \subsection{Additional Readings}
    \begin{readinglist}
        \item \bibentry{kuznetsov_motivations_2006}
        \item \bibentry{lakhani_why_2003}
        \item \bibentry{lampe_motivations_2010}
        \item \bibentry{roberts_understanding_2006}
        \item \bibentry{rafaeli_online_2008}
        \item \bibentry{yang_motivations_2010}
    \end{readinglist}

\section{3 -- Commitment}
\textcolor{CUGold}{\textbf{Wednesday, August 31}}\\
Why do participants stick around and become active contributors within online communities? We discuss how the design impacts user commitment as well as a framework explaining how users transition into leadership roles.

    \subsection{Readings}
    \begin{readinglist}
        \item \bibentry{wikipedia_wikifauna_2016}
        \item \bibentry{kraut_building_2012}. Chapter 3.
        \item \bibentry{preece_readerleader_2009}
    \end{readinglist}
    
    \subsection{Assignments}
    \begin{description}%[itemsep=2pt,labelsep=0pt]
        \item[Reading Response 2 ] -- due Wednesday, August 31. What kinds of motivations are particular to writing an encyclopedia and what motivations might generalize across contexts? What motivations are most likely to sustain users' commitment over the long term? What are potential challenges when a system needs to support diverse motivations? What are the risks of systems that prioritize some motivations over others?
    \end{description}
    
    \subsection{Additional Readings}
    \begin{readinglist}
        \item \bibentry{panciera_wikipedians_2009}.
        %\item \bibentry{}
        item \bibentry{}
    \end{readinglist}

%%%%%%%%%%%%%
%% WEEK 03 %%
%%%%%%%%%%%%%
\section{4 -- Labor Day}
\textcolor{CUGold}{\textbf{Monday, September 5 --- NO CLASS}}\\

\section{5 -- Newcomers}
\textcolor{CUGold}{\textbf{Wednesday, September 7}}\\
How do peer production systems handle new users? What kinds of structures discourage newcomers? We will discuss and evaluate social and technical interventions to support newcomers.

    \subsection{Readings}
    \begin{readinglist}
        \item \bibentry{wikipedia_seven_2015}
        \item \bibentry{halfaker_dont_2011}
        \item \bibentry{morgan_tea_2013}
    \end{readinglist}

%%%%%%%%%%%%%
%% WEEK 04 %%
%%%%%%%%%%%%%
\section{6 -- Culture and Norms}
\textcolor{CUGold}{\textbf{Monday, September 12}}\\
What does it mean to become a member of a peer production community? What kinds of behaviors are encouraged? We will discuss the different norms that shape community behavior.

    \subsection{Readings}
    \begin{readinglist}
        \item \bibentry{wikiphilosophies}
        %\item Reagle, J. (2010). Good faith collaboration : the culture of Wikipedia. MIT Press, Cambridge Mass. Chapters 1 \& 3.
        \item \bibentry{bryant_becoming_2005}
        \item \bibentry{kriplean_articulations_2008}
    \end{readinglist}

    \subsection{Additional Readings}
    \begin{readinglist}
        \item \bibentry{burke_mopping_2008}
        \item \bibentry{danescu-niculescu-mizil_echoes_2012}
        \item \bibentry{thom_whats_2009}
        \item \bibentry{zhu_organizing_2012}
    \end{readinglist}

\section{7 -- Rules and Governance}
\textcolor{CUGold}{\textbf{Wednesday, September 14}}\\
In addition to norms, communities also have more formal governance structures like written rules and practices for elites to use power. We will discuss the emergence and consequences of these rules and governance mechanisms for the sustainability of the community.

    \subsection{Readings}
    \begin{readinglist}
        \item \bibentry{butler_dont_2008}
        \item \bibentry{shaw_laboratories_2014}
    \end{readinglist}
    
    \subsection{Assignments}
    \begin{description}%[itemsep=2pt,labelsep=0pt]
        \item[Reading Response 3 ] -- due Wednesday, September 14. Are the various wiki-philosophies specific to a system like Wikipedia or could they extend to other peer production contexts? What are the trade-offs to consider when adopting more formal rules? What strategies could peer production systems implement to deal with the Iron Law?
    \end{description}
    
    \subsection{Additional Readings}
    \begin{readinglist}
        \item \bibentry{beschastnikh_wikipedian_2008}
        \item \bibentry{black_self-governance_2011}
        \item \bibentry{forte_decentralization_2009}
        \item \bibentry{kriplean_community_2007}
        \item \bibentry{kostakis_peer_2010}
        %\item \bibentry{konieczny_governance_2009}
        \item \bibentry{konieczny_adhocratic_2010}
        %\item \bibentry{leskovec_governance_2010}
        \item \bibentry{omahony_emergence_2007}
    \end{readinglist}


%%%%%%%%%%%%%
%% WEEK 05 %%
%%%%%%%%%%%%%
\section{8 -- Conflict}
\textcolor{CUGold}{\textbf{Monday, September 19}}\\
Conflict is inevitable as social groups grow larger. What kinds of conflict in peer production systems and how does the community manage them? We will discuss how Wikipedia manages conflicts through a several technical and social processes.

    \subsection{Readings}
    \begin{readinglist}
        \item \bibentry{kittur_he_2007}
        \item \bibentry{joyce_handling_2011}
    \end{readinglist}
    
    \subsection{Additional Readings}
    \begin{readinglist}
        \item \bibentry{arazy_stay_2013}
        \item \bibentry{borra_societal_2015}
        \item \bibentry{kittur_beyond_2010}
        \item \bibentry{matei_wikipedias_2010}
        \item \bibentry{osman_role_2013}
        \item \bibentry{sumi_edit_2011}
        \item \bibentry{towne_your_2013}
        \item \bibentry{yasseri_dynamics_2012}
    \end{readinglist}

\section{9 -- Bias}
\textcolor{CUGold}{\textbf{Wednesday, September 21}}\\
How susceptible are peer production systems to the biases of their contributors? We will discuss research findings about geographic, cultural, gender, and political biases in Wikipedia.

    \subsection{Readings}
    \begin{readinglist}
        \item \bibentry{hecht_measuring_2009}
        \item \bibentry{wagner_women_2016}
        \item \bibentry{greenstein_is_2012}
    \end{readinglist}
    
    \subsection{Assignments}
    \begin{description}%[itemsep=2pt,labelsep=0pt]
        \item[Reading Response 4 ] -- due Wednesday, September 21. How do conflict resolution strategies need to change as peer production communities change in size, complexity, or mission? What are trade-offs to consider around conflict management when bringing more diverse contributors into the system? What are other kinds of bias that can creep into peer production systems?
    \end{description}
    
    \subsection{Additional Readings}
    \begin{readinglist}
        \item \bibentry{antin_gender_2011}
        \item \bibentry{collier_conflict_2012}
        \item \bibentry{graellsgarrido_first_2015}
        \item \bibentry{iosub_emotions_2014}
        \item \bibentry{johnson_not_2016}
        \item \bibentry{lam_wpclubhouse_2011}
        \item \bibentry{reagle_gender_2011}
    \end{readinglist}

%%%%%%%%%%%%%
%% WEEK 06 %%
%%%%%%%%%%%%%
\section{10 -- Revision Histories}
\textcolor{CUGold}{\textbf{Monday, September 26}}\\
Wikipedia records every change made to every article since its inception. We will discuss how to access these data and what is recorded. We will look at two papers that used these data to understand the dynamics of conflict between users and well as how to measure how much time users invest in editing Wikipedia.

    \subsection{Readings}
    \begin{readinglist}
        \item \bibentry{viegas_studying_2004}
        \item \bibentry{geiger_using_2013}
    \end{readinglist}
    
    \subsection{Additional Readings}
    \begin{readinglist}
        \item \bibentry{ekstrand_rv_2009}
    \end{readinglist}

\section{11 -- Revision Histories Lab}
\textcolor{CUGold}{\textbf{Wednesday, September 28}}\\
We will share and discuss the findings from our laboratory assignments.

    \subsection{Assignments}
    \begin{description}%[itemsep=2pt,labelsep=0pt]
        \item[Lab Assignment 1 ] -- due September 28. 
    \end{description}

%%%%%%%%%%%%%
%% WEEK 07 %%
%%%%%%%%%%%%%
\section{12 -- Networks}
\textcolor{CUGold}{\textbf{Monday, October 3}}\\
Wikipedia encodes a variety of relationships in its data. We will discuss these different kinds of relationships and explore how to use methods from network science to understand and visualize these data. 

    \subsection{Readings}
    \begin{readinglist}
        \item \bibentry{keegan_analyzing_2013}
        \item \bibentry{kane_its_2009}
    \end{readinglist}
    
    \subsection{Additional Readings}
    \begin{readinglist}
        \item \bibentry{aragon_biographical_2012}
        \item \bibentry{brandes_network_2009}
        %\item \bibentry{capocci_preferential_2006}
        %\item \bibentry{howison_validity_2011}
        \item \bibentry{jankowski-lorek_verifying_2016}
        \item \bibentry{keegan_editors_2012}
        \item \bibentry{laniado_when_2011}
        %\item \bibentry{laniado_co-authorship_2011}
        \item \bibentry{massa_social_2011}
        \item \bibentry{suh_us_2007}
        \item \bibentry{welser_finding_2011}
    \end{readinglist}

\section{13 -- Networks Lab}
\textcolor{CUGold}{\textbf{Wednesday, October 5}}\\
We will share and discuss the findings from our laboratory assignments.

    \subsection{Assignments}
    \begin{description}%[itemsep=2pt,labelsep=0pt]
        \item[Lab Assignment 2 ] -- due Wednesday, October 5. 
    \end{description}

%%%%%%%%%%%%%
%% WEEK 08 %%
%%%%%%%%%%%%%
\section{14 -- Pageviews}
\textcolor{CUGold}{\textbf{Monday, October 10}}\\
Wikipedia collects data about the volume of pageviews for each article. What do these patterns of information consumption reveal? We will explore how they be combined with other kinds of data to measure inefficiencies or make forecasts.

    \subsection{Readings}
    \begin{readinglist}
        \item \bibentry{generous_global_2014}
        \item \bibentry{warncke-wang_misalignment_2015}
    \end{readinglist}
    
    \subsection{Additional Readings}
    \begin{readinglist}
        \item \bibentry{althoff_analysis_2013}
        \item \bibentry{choi_predicting_2012}
        \item \bibentry{crane_robust_2008}
        \item \bibentry{hickmann_forecasting_2015}
        \item \bibentry{kampf_fluctuations_2012}
        \item \bibentry{mestyan_early_2013}
        \item \bibentry{moat_quantifying_2013}
        \item \bibentry{ratkiewicz_characterizing_2010}
        \item \bibentry{thij_modeling_2012}
    \end{readinglist}

\section{15 -- Pageviews Lab}
\textcolor{CUGold}{\textbf{Wednesday, October 12}}\\
We will share and discuss the findings from our laboratory assignments.

    \subsection{Assignments}
    \begin{description}%[itemsep=2pt,labelsep=0pt]
        \item[Lab Assignment 3 ] -- due Wednesday, October 12. 
    \end{description}
    
%%%%%%%%%%%%%
%% WEEK 09 %%
%%%%%%%%%%%%%
\section{16 -- Ontologies and Linked Data}
\textcolor{CUGold}{\textbf{Monday, October 17}}\\
Wikipedia encodes a variety of semantic and other structured data in infoboxes, categories, templates, and other structures. How are these data structured and linked together? What are the opportunities and challenges of new projects like WikiData?

    \subsection{Readings}
    \begin{readinglist}
        \item \bibentry{shirky_ontology_2005}
        \item \bibentry{vrandecic_wikidata_2014}
    \end{readinglist}
    
    \subsection{Additional Readings}
    \begin{readinglist}
        \item \bibentry{bizer_emerging_2009}
        \item \bibentry{erxleben_introducing_2014}
        \item \bibentry{kittur_whats_2009}
        \item \bibentry{volkel_semantic_2006}
        \item \bibentry{yu_ontology_2007}
    \end{readinglist}

\section{17 -- Ontologies Lab}
\textcolor{CUGold}{\textbf{Wednesday, October 19}}\\
We will share and discuss the findings from our laboratory assignments.

    \subsection{Assignments}
    \begin{description}%[itemsep=2pt,labelsep=0pt]
        \item[Lab Assignment 4 ] -- due Wednesday, October 19. 
    \end{description}
    
%%%%%%%%%%%%%
%% WEEK 10 %%
%%%%%%%%%%%%%
\section{18 -- Bots}
\textcolor{CUGold}{\textbf{Monday, October 24}}\\
Peer production systems require repetitive maintenance tasks that lend themselves to automation. What kinds of tasks can be automated? What are the consequences of relying on bots to help maintain the system?

    \subsection{Readings}
    \begin{readinglist}
        \item \bibentry{geiger_work_2010}
        \item \bibentry{geiger_when_2013}
    \end{readinglist}
    
    \subsection{Additional Readings}
    \begin{readinglist}
        \item \bibentry{halfaker_snuggle_2014}
        \item \bibentry{muller-birn_work_2013}
        \item \bibentry{steiner_bots_2014}
    \end{readinglist}

\section{19 -- Bots Lab}
\textcolor{CUGold}{\textbf{Wednesday, October 26}}\\
We will share and discuss the findings from our laboratory assignments.

    \subsection{Assignments}
    \begin{description}%[itemsep=2pt,labelsep=0pt]
        \item[Lab Assignment 5 ] -- due Wednesday, October 26. 
    \end{description}
    
%%%%%%%%%%%%%
%% WEEK 11 %%
%%%%%%%%%%%%%
\section{20 -- Edge Case 1: High-Tempo Collaboration}
\textcolor{CUGold}{\textbf{Monday, October 31}}\\
Wikipedia's social and technical structures were organized assuming that its knowledge would be established and stable. However, Wikipedians have demonstrated a remarkable capacity to rapidly create and revise articles in response to current and breaking news events. 

    \subsection{Readings}
    \begin{readinglist}
        \item \bibentry{keegan_hot_2011}
        \item \bibentry{keegan_history_2013}
    \end{readinglist}
    
    \subsection{Additional Readings}
    \begin{readinglist}
        %\item \bibentry{bechky_gaffers_2006}
        \item \bibentry{bechky_expecting_2011}
        \item \bibentry{bigley_incident_2001}
        \item \bibentry{faraj_coordination_2006}
        \item \bibentry{keegan_is_2015}
        \item \bibentry{klein_dynamic_2006}
        \item \bibentry{majchrzak_coordinating_2007}
        \item \bibentry{palen_emergence_2008}
        \item \bibentry{palen_crisis_2016}
        %\item \bibentry{weick_collective_1993}
        \item \bibentry{weick_organizing_1999}
    \end{readinglist}

\section{21 -- Edge Case 2: Alternative Internet Encyclopedias}
\textcolor{CUGold}{\textbf{Wednesday, November 2}}\\
The MediaWiki software and encyclopedic mission of Wikipedia were adopted by other online communities that have very different norms and values. We will discuss how these alternative norms and values influenced the success and sustainability of these encyclopedias.

    \subsection{Readings}
    \textbf{\textit{Note}: These alternative encyclopedias contain content that you may find offensive.}\vspace{-10pt}
    \begin{readinglist}
        \item \bibentry{colbert_schlafly_2009}
        \item \bibentry{sanger_why_2007}
        \item \bibentry{creationwiki_about}
        \item \bibentry{rationalwiki_about}
        \item \bibentry{conservapedia_about}
        \item \bibentry{dramatica_about}
    \end{readinglist}
    
    \subsection{Assignments}
    \begin{description}%[itemsep=2pt,labelsep=0pt]
        \item[Reading Response 5 ] -- due Wednesday, November 2. What shortcomings did the creators of these alternative encyclopedias identify with Wikipedia's model? How were their alternative values implemented in the code of the system and practices of the community? How should we judge the success of their alternative approaches?
    \end{description}
    
    \subsection{Additional Readings}
    \begin{readinglist}
        \item \bibentry{lam_past_2011}
        \item \bibentry{hill_almost_2013}
        \item \bibentry{oneil_shirky_2010}
        \item \bibentry{sanger_fate_2009}
        \item \bibentry{shirky_larry_2006}
        \item \bibentry{sundin_debating_2007}
    \end{readinglist}
    
%%%%%%%%%%%%%
%% WEEK 12 %%
%%%%%%%%%%%%%
\section{22 -- Successors 1: Citizen Science}
\textcolor{CUGold}{\textbf{Monday, November 7}}
Wikipedia demonstrated that a long tail of non-experts can produce a large and high-quality encyclopedia. How effective are aggregating the contributions from other non-experts at solving other scientific tasks? We will explore how lessons from peer production have influenced the development of citizen science.

    \subsection{Readings}
    \begin{readinglist}
        \item \bibentry{rotman_dynamic_2012}
        \item \bibentry{wiggins_conservation_2011}
    \end{readinglist}
    
    \subsection{Additional Readings}
    \begin{readinglist}
        \item \bibentry{anderson_setihome_2002}
        \item \bibentry{cooper_predicting_2010}
        \item \bibentry{lintott_galaxy_2008}
        \item \bibentry{reed_exploratory_2013}
        \item \bibentry{sakaki_earthquake_2010}
        \item \bibentry{stefanovitch_error_2014}
        \item \bibentry{yuen_survey_2011}
    \end{readinglist}
    

\section{23 -- Successors 2: OpenStreetMaps}
\textcolor{CUGold}{\textbf{Wednesday, November 9}}\\
The peer production model has also been adopted to produce other common goods such as maps. We will discuss how ``geo-wikis'' work and what additional considerations are needed when moving away from textual data. \textbf{Featuring a guest lecture by Jennings Anderson}.

    \subsection{Readings}
    \begin{readinglist}
        \item \bibentry{haklay_openstreetmap_2008}
        \item \bibentry{priedhorsky_computational_2008}
    \end{readinglist}
    
    \subsection{Assignments}
    \begin{description}%[itemsep=2pt,labelsep=0pt]
        \item[Reading Response 6 ] -- due Wednesday, November 9. TBD.
    \end{description}
    
    \subsection{Additional Readings}
    \begin{readinglist}
        \item \bibentry{budhathoki_motivation_2013}
        \item \bibentry{elwood_researching_2012}
        \item \bibentry{goodchild_citizens_2007}
        \item \bibentry{mooney_towards_2010}
        \item \bibentry{palen_success_2015}
        \item \bibentry{quattrone_exploring_2016}
        %\item \bibentry{soden_infrastructure_2016}
        \item \bibentry{wang_crowdatlas_2013}
    \end{readinglist}
    
%%%%%%%%%%%%%
%% WEEK 13 %%
%%%%%%%%%%%%%
\section{24 -- Creative Collaboration 1: Scratch}
\textcolor{CUGold}{\textbf{Monday, November 14}}\\
Peer production communities can also collaborate to produce creative work. Scratch is an online community of young people who create, share, and remix animations, games, and other computational projects. \textbf{Featuring a guest lecture by Professor Ricarose Roque}.

    \subsection{Readings}
    \begin{readinglist}
        \item \bibentry{resnick_scratch_2009}
        \item \bibentry{roque_supporting_2016}
    \end{readinglist}
    
    \subsection{Additional Readings}
    \begin{readinglist}
        \item \bibentry{dasgupta_remixing_2016}
        \item \bibentry{hill_cost_2013}
        \item \bibentry{monroy-hernandez_computers_2011}
        \item \bibentry{nickerson_appropriation_2011}
    \end{readinglist}

\section{25 -- Creative Collaboration 2: Threadless}
\textcolor{CUGold}{\textbf{Wednesday, November 16}}\\
How can crowds be utilized within traditional organizations to support research and development? We will examine the case of the Threadless community and how the crowd is used to generate and evaluate creative designs.

    \subsection{Readings}
    \begin{readinglist}
        \item \bibentry{brabham_moving_2010}
        \item \bibentry{kittur_crowdsourcing_2010}
    \end{readinglist}
        
    \subsection{Assignments}
    \begin{description}%[itemsep=2pt,labelsep=0pt]
        \item[Reading Response 7 ] -- due Wednesday, November 16. TBD.
    \end{description}
    
    \subsection{Additional Readings}
    \begin{readinglist}
        \item \bibentry{brabham_crowdsourcing_2008}
        \item \bibentry{brabham_moving_2010}
        \item \bibentry{geiger_managing_2011}
        \item \bibentry{hutter_communitition_2011}
        \item \bibentry{yu_cooks_2011}
        \item \bibentry{yu_collective_2012}
    \end{readinglist}
    
%%%%%%%%%%%%%
%% WEEK 14 %%
%%%%%%%%%%%%%
\section{26 -- Fall Break}
\textcolor{CUGold}{\textbf{Monday, November 21  --- NO CLASS}}

\section{27 -- Fall Break}
\textcolor{CUGold}{\textbf{Wednesday, November 23 --- NO CLASS}}

%%%%%%%%%%%%%
%% WEEK 15 %%
%%%%%%%%%%%%%
\section{28 -- Beyond Wikis: Open Hardware}
\textcolor{CUGold}{\textbf{Monday, November 28}}\\
Peer production models can also extend beyond online and software collaborations. What strategies have members of the Open Hardware movement translated from elsewhere and what have they had to innovate themselves?

    \subsection{Readings}
    \begin{readinglist}
        \item \bibentry{buechley_lilypad_2010}
        \item \bibentry{kuznetsov_rise_2010}
        \item \bibentry{weiss_open_2008}
    \end{readinglist}
    
    \subsection{Additional Readings}
    \begin{readinglist}
        \item \bibentry{buechley_diy_2009}
        \item \bibentry{malinen_open_2010}
        \item \bibentry{mellis_fab_2011}
        \item \bibentry{mellis_collaboration_2012}
    \end{readinglist}

\section{29 -- Beyond Wikis: Blockchains}
\textcolor{CUGold}{\textbf{Wednesday, November 30}}\\
Peer production can also be used to ensure accountability across a complex system. Crypto-currencies like Bitcoin introduced the idea of ``blockchains'' that offer some radical ideas about how to decentralize control over important transactions.

    \subsection{Readings}
    \begin{readinglist}
        \item \bibentry{swan_blockchain_2015}. Skim Chapters 1--3.
    \end{readinglist}
    
    \subsection{Assignments}
    \begin{description}%[itemsep=2pt,labelsep=0pt]
        \item[Reading Response 8 ] -- due Wednesday, November 30. TBD.
    \end{description}
    
%%%%%%%%%%%%%
%% WEEK 16 %%
%%%%%%%%%%%%%
\section{30 -- Platform Cooperativism}
\textcolor{CUGold}{\textbf{Monday, December 5}}\\
Peer production demonstrated the potential for information technology to support the production of valuable goods in the absence of traditional economic incentives. How far could these models be applied to other areas of the economy? \textbf{Featuring a guest lecture by Professor Nathan Schneider.}

    \subsection{Readings}
    \begin{readinglist}
        \item \bibentry{louis_user_2013}
        \item \bibentry{scholz_platform_2014}
        %\item \bibentry{schneider_internet_2015}
    \end{readinglist}

\section{31 -- Wrap-up}
\textcolor{CUGold}{\textbf{Wednesday, December 7}}\\
Discuss final project preparations, reflect on themes from the course, overflow time if there are topics we missed or want to dive into greater depth.


%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% END OF SEMESTER %%%%
%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%

% bibliography here
\newpage
\renewcommand{\bibsection}{\section{\huge \bibname}\prebibhook}
\baselineskip 14.2pt
\nobibliography{refs}
\bibliographystyle{apalike}

\end{document}
